---
title: "HW 11: Data Visualization I"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
#### Load libraries

```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```


***
#### Habitat parameters for Aegla aegla_mutated


*Import Data*

```{r}
aegla_data <- read_csv(here("data", "aegla_crabs.csv"))
```


*wrangle Data*

Use select() to remove the Site, AT, WT, Crab, Shrimp, Aeglam, AeglaFe, AeglaFo columns.

Use drop_na() to remove the one row with missing water quality data.

Create a row_order dummy variable like we did for the aegla_mutated.

You must use the pipe (%>%) to write efficient code.

```{r}
(aegla_mutated <- aegla_data %>%
  select(!c(Site, AT, WT, Crab, Shrimp, Aeglam, AeglaFe, AeglaFo)) %>%
  drop_na() %>%
  mutate(row_order = 1:nrow(.)))
```


*Graph Data*

**Explore the data. There are three apparent outliers among the variables but only one that I think is an actual outlier. Find them and justify your choices.***



```{r}
p_width <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Width)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_depth <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Depth)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_flow <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_ph <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = pH)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_tds <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = TDS)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_cond <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Cond)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_n <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = N)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_ni <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Ni)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())


p_nt <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Nt)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_phos <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Phos)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_mg <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Mg)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_ca <- aegla_mutated %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = Ca)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_width + p_depth + p_flow +
  p_ph + p_tds + p_cond +
  p_n + p_ni + p_nt +
  p_phos + p_mg + p_ca +
  plot_layout(ncol = 3)

```

The three apparent outliers are - Depth, pH and Phos. Though all three have a point that is clearly away from the other observations, data of Depth and pH is much more scattered and distance away from main data is less compared to Phos. All observations of Phos are less than 0.25 and the outlier lies above 0.75. This difference is very vast hence Phos is the one with the actual outlier. 

*More on above statement*

Outlier is suppose to be incorrectly recorded data. By definition it lies beyond 1.5 IQR. So lets make boxplots-

```{r}
aegla_box <- aegla_mutated %>%
  mutate(box_const = "boxplot")

  ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Width))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Depth))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Flow))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = pH))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = TDS))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Cond))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = N))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Ni))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Nt))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Phos))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Mg))

ggplot(aegla_box,aes(x = box_const)) +
  geom_boxplot(aes(y = Ca))

```

TDS, Cond, Ni, Mg and Ca are the only ones that do not have an outlier by definition.


```{r}
p_width + p_depth + p_flow +
  p_ph + p_n + p_nt +
  p_phos
  
```

Visually, in a data of 55 rows one cannot call 4-8 points an error under a variable. This especially applies for Flow and Nt.

Under width, the point at 11.3 is separate from its next closest by 2 grid (unit in plot) units. This may be classified under outlier but would not change the data much.

Under Depth and pH the majority of data is scattered in 5 grid lengths. The outlier in Depth has a distance of 5 grids from its closest neighbour. This does appear to be an outlier. In case of pH most of the data is concentrated on 7.0. Visually the the outlier at 6.0 stands out, only because there are not enough scatter below 7.0. This also appears to be an outlier.

In Phos most of the data lies in 1 grid unit and the outlier is 6 grid units away from its closest neighbour. This is definitely an outlier.

In N the scatter is is gradual and does not stand out, so not an outlier.



**Choose four variables to make Cleveland plots, and make a 2x2 grid with the four plots, using patchwork.**

```{r}
p1 <- ggplot(aegla_mutated, aes(y = row_order)) +
  geom_point(aes(x = Flow)) +
  labs(y = "Order of data in the dataset")

p2 <- ggplot(aegla_mutated, aes(y = row_order)) +
  geom_point(aes(x = pH)) +
  labs(y = "Order of data in the dataset")

p3 <- ggplot(aegla_mutated, aes(y = row_order)) +
  geom_point(aes(x = Nt)) +
  labs(x = "Nitrite concentrations",
       y = "Order of data in the dataset")

p4 <- ggplot(aegla_mutated, aes(y = row_order)) +
  geom_point(aes(x = Phos)) +
  labs(x = "Phosphate concentrations",
       y = "Order of data in the dataset")

p1 + p2 + p3 + p4 + plot_layout(ncol = 2)
```

From the plots one can deduce- 
- Most of observations in flow lie between 0.3 and 0.5, further within that most lie on 0.4.
- pH of observations is usually neutral, sometimes slightly basic and very rarely acidic.  
- Most observations lie on two readings- 0.000 and 0.001. Most of the data collected is from places with no nitrite concentration.
- All except one phosphate concentration lie between 0.00 and 0.1875.


**Choose three different variables to make histograms. Play with the bins or binwidths argument that you think reveals the data well. Hint: the default bins = 30 that geom_histogram is usually not a suitable choice. Save each histogram to a unique variable.**

```{r}
(h_mg <- ggplot(aegla_mutated) +
  geom_histogram(aes(x = Mg),
                 binwidth = 0.1,
                 boundary = 0,
                 closed = "right",
                 color = "black"))
```
The values appear to be in clusters and there are empty bins near tall bars. There are 10 bars between a unit. The tallest bar with 0.1 width is 3.51 to 3.60.

```{r}
(h_ca <- ggplot(aegla_mutated) +
  geom_histogram(aes(x = Ca),
                 binwidth = 0.1,
                 boundary = 0,
                 closed = "right",
                 color = "black"))
```

Here the tallest bar is 0.00-0.10. The data does not follow a particular type of distribution. The rest of the bars are of random lengths (no gradual increase in length then gradual decrease).

```{r}
(h_width <- ggplot(aegla_mutated) +
  geom_histogram(aes(x = Width),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black"))
```
The tallest bar is 5.51 - 6.0 (can also be seen as densely packed in p_width). Here data seems to me scattered in the range 1 - 4.

**Use the same three variables to produce density plots. Save each plot to a unique variable.**

```{r}
(d_mg <- ggplot(aegla_mutated) +
  geom_density(aes(x = Mg)))

(d_ca <- ggplot(aegla_mutated) +
  geom_density(aes(x = Ca)))

(d_width <- ggplot(aegla_mutated) +
  geom_density(aes(x = Width)))
```

From the above plots it may be noted that the graph shows some value of density even though at an exact point there may not be any observations present. 

In Mg the graph increases from 0.1 to 0.25 then decreases again to 0.1. It then seems to increase gradually then decrease suddenly. Overall, it seems to increase.

In Ca the graph increases (very little) then gradually decreases (for a long range) the increases(very slowly) then decreases again. Overall, it seems to decrease.

In Width the graph seems to have 2 peaks, the third is just sudden decrease in decrease of density.



**Use `patchwork to produce a 2 column matrix with the histograms in the left column and the corresponding density plots in the right column.**

```{r}
(h_mg | d_mg) /
(h_ca | d_ca) /
(h_width | d_width)
```

The peaks in density plots correspond more to clusters of bars (of considerable length) and not always the height of the bars. This can be explained as height and size of bars were decided manually and the height changes based of binwidth decided. 


**Use ggpairs to explore the relationships among all the variables.**

```{r}
aegla_mutated %>% ggscatmat(columns = 1:12)
```

Ca concentration and conductivity have a strong negative correlation (-0.73). Ca and TDS also have a strong negative correlation (-0.74).
Mg has strong positive correlation with Cond and TDS (0.76 and 0.77 respectively). Ca and Mg have a correlation of (-0.69).

### END
