---
title: "R notebook HW03"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Homework 03: R Markdown

Part 1: An R Notebook case study



Load the packages.

```{r}
library(tidyverse)
library(knitr)
library(xtable)
```



Special data set- anscombe viewed

```{r}
anscombe
```



Wrangle the data
```{r}
ansc <- 
  bind_cols(
    anscombe %>% 
      select(matches("x")) %>% 
      gather(dataset, "x"),
    anscombe %>% 
      select(matches("y")) %>% 
      gather(key,"y")
  ) %>% 
  select(-key) %>% 
  mutate(dataset = gsub("x","", dataset)) %>% 
  as_tibble()
  
```



View the wrangled data set
```{r}
ansc
```



Performing some calculations
```{r}
ansc %>% 
  group_by(dataset) %>% 
  summarize(
    mean_x = mean(x),
    stdev_x = sd(x),
    mean_y = mean(y),
    stdev_y = sd(y)
  )
```



Visualize the Data
```{r}
ansc %>% ggplot(aes(x, y, color = dataset)) + geom_point() + facet_wrap(~dataset, ncol = 2)
```




Part 2: Two more R Markdown tricks



View first few rows of iris
```{r}
head(iris)
```



Using kable to align and name the columns
```{r}
kable(head(iris),
      align = c("l", "c", "r", "c", "l"),
      col.names = c("Sepal Length", "Sepal Width", "Sepal Length", "Sepal Width", "Species"))
```



View first few rows of airquality
```{r}
head(airquality)
```



Using kable to align and rename the columns
```{r}
kable(head(airquality),
      align = c("r", "r", "c", "c", "l", "l"),
      col.names = c("Ozone (ppb)", "Solar R (lang)", "Wind (mph)", "Temp (°F)", "Month", "Day"))
```



Quick statistical test to see if there is a significant relationship between weight and feed
```{r}
chicks_lm <- lm(weight ~ feed, data = chickwts)
summary(chicks_lm)
```



Insert a code chunk and run kable() on chicks_lm. What happens? Write a 1-2 sentence summary of the results.[^Footnote]
```{r error=TRUE}
kable(chicks_lm)
```
In the above code an error message comes stating- "Error in as.data.frame.default(x) : cannot coerce class ‘"lm"’ to a data.frame". Kable runs on data in the form of a data frame with rows and columns, the variable- chicks_lm- is not in such format.



Now try it-
Use lm on the trees dataset to see if Girth is a function of Height. Assign the results to a suitably-named variable.
Use xtable() and kable() to produce a table with the results.
```{r}
trees_lm <- lm(Girth ~ Height, data = trees)
xtable(trees_lm) %>% kable()
```

Equations 
The Chi-square equation
$$\chi_c^2 = \Sigma\frac{(O_i - E_i)^2}{E_i
}$$

[^Footnote]: Error code included.



