---
title: "HW 08: More Data Wrangling"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

**You do not answer any of the questions this week.** Just read and run all of the examples as noted below. Pay attention to the code you enter for the examples. You will be applying the information in future assignments.


Require packages

```{r}
library(tidyverse)
library(nycflights13)
```


***
### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 


create a narrower dataset and add columns using mutate()

```{r}
flights_sml <- select(flights,
                      year:day,
                      ends_with("delay"),
                      distance,
                      air_time
                      )
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
       )
```

Add more columns using columns just created.

```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )
```

Keep only the new variables

```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours
          )
```


*5.5.1 Useful creation functions*

Modular arithmetic- hour and minute from dep_time

```{r}
transmute(flights, 
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100
          )
```

Offsets: using lead() and lag()

```{r}
(x <- 1:10)
lag(x)
lead(x)
```


############# Problematic
Cumulative and rolling aggregates

```{r}
x 
cumsum(x)
cummean(x)
```

##########################

Rank the numbers in a variable

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

Other variants

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```


***
### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.


*5.6 Grouped summaries with summarise()*

Collapses a data frame to a single row using summarise()

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

Get the average delay per date

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

*5.6.1 Combining multiple operations with the pipe*


























***
### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.


































