---
title: "HW 08: More Data Wrangling"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

**You do not answer any of the questions this week.** Just read and run all of the examples as noted below. Pay attention to the code you enter for the examples. You will be applying the information in future assignments.


Require packages

```{r}
library(tidyverse)
library(nycflights13)
```


***
### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 


create a narrower dataset and add columns using mutate()

```{r}
flights_sml <- select(flights,
                      year:day,
                      ends_with("delay"),
                      distance,
                      air_time
                      )
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
       )
```

Add more columns using columns just created.

```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )
```

Keep only the new variables

```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours
          )
```


*5.5.1 Useful creation functions*

Modular arithmetic- hour and minute from dep_time

```{r}
transmute(flights, 
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100
          )
```

Offsets: using lead() and lag()

```{r}
(x <- 1:10)
lag(x)
lead(x)
```


Cumulative and rolling aggregates

```{r}
x 
cumsum(x)
cummean(x)
```
Note- Results of cummean() unexpected (do not match book) for unknown reasons.

Rank the numbers in a variable

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

Other variants

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```


***
### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.


*5.6 Grouped summaries with summarise()*

Collapses a data frame to a single row using summarise()

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

Get the average delay per date

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

*5.6.1 Combining multiple operations with the pipe*

Relationship between the distance and average delay for each location

```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)
                   )
delay <- filter(delay, count > 20, dest != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

Do the above using pipes

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

*5.6.2 Missing values*

Summarise using na.rm = FALSE (the default)

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

Remove the missing values prior to computation

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

Remove cancelled flights first then make tibble as above

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```


*5.6.3 Counts*

Planes (identified by their tail number) with the highest average delays

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) +
  geom_freqpoly(binwidth = 10)
```

Draw a scatterplot of number of flights vs. average delay

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```

Filter out the groups with the smallest numbers of observations

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```

**Skipped the baseball part of 5.6.3**

*5.6.4 Useful summary functions*

Find average and average positive delay in arr_delay

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0])
  )
```

Why is distance to some destinations more variable than to others? Standard deviation of distance grouped by destination.

```{r}
not_cancelled %>% 
group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

When do the first and last flights leave each day?

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```

First and last departure for each day

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time),
    last_dep = last(dep_time)
  )
```

Filter of rank

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

Which destinations have the most carriers?

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

Count the number of each destination (how often each reoccurs)

```{r}
not_cancelled %>% 
  count(dest)
```

“count” (sum) the total number of miles each plane flew

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

How many flights left before 5am?

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))
```

What proportion of flights are delayed by more than an hour?

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_prop = mean(arr_delay > 60))
```


*5.6.5 Grouping by multiple variables*

Find number of flights per day, per month and per year.

```{r}
daily <- group_by(flights, year, month, day)
(per_day <- summarise(daily, flights = n()))

(per_month <- summarise(per_day, flights = sum(flights)))

(per_year <- summarise(per_month, flights = sum(flights)))
```


*5.6.6 Ungrouping*

Remove grouping, and return to operations on ungrouped data

```{r}
daily %>% 
  ungroup() %>% 
  summarise(flights = n())
```



***
### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

Find the worst members of each group

```{r}
flights_sml %>% 
  group_by(year, month, day) %>% 
  filter(rank(desc(arr_delay)) < 10)
```


Find all groups bigger than a threshold

```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```

Standardise to compute per group metrics

```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```


###################

