---
title: 'HW07: Data Importing and Wrangling 1'
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Load the packages and file path.

```{r}
library(tidyverse)
library(here)
```


#### Limpets

Requirements:

Data file: limpets.csv.

Are the data tidy?
Yes, all three columns are unique variables- Length, Site, Protected- each observation is a row.

Use the col_types argument. The three column types are numeric, character, and character.

```{r}
limpets_data <- read_csv(here("data", "limpets.csv"), col_types = "ncc")
```

Make the Sites column an ordered factor with these levels:
PBL, KNRM, VBG, WP, PF, DP, SIO, CTZ1, CTZ2, CNM

```{r}
limpets_data$Site <- factor(limpets_data$Site, ordered = TRUE, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"))
```

PBL is the northern most site (Pebble Beach). CNM is the southernmost site (Cabrillo National Monument).

Make a boxplot of length (in millimeters) for each site, colored by protected status.

Change the axis labels so they begin with capital letters. The y-axis should include the unit of measurement in parentheses.

Note: Both axes labels already begin with capital letters. 


```{r}
limpets_data %>% 
  ggplot() +
  geom_boxplot(aes(x = Site, y = Length, color = Protected)) +
  labs(y = "Length(in millimeters)")
```

In your description, tell which two sites have outliers and whether the protected sites tend to have larger or smaller limpets.

The two sites that have outliers are WP and DP. Protected sites tend to have larger limpets than unprotected ones.






##########################




#### Roseate Terns

Data: roseate_terns.txt

```{r}
roseate_data <- read_tsv(here("data", "roseate_terns.txt"))
```

Are the data tidy?
No, the sites are individual columns rather than being under one variable of "Sites".

```{r}
(roseate_tidy <- roseate_data %>% 
  pivot_longer(c(AR, AL, AC, BR, BL,BC, JR, JL, JC ), names_to = "Sites", values_to = "Population"))
roseate_tidy$Sites <- factor(roseate_tidy$Sites)
```

Variables "Population" and "Sites" assigned. Data tidied. New data is 'roseate_tidy' 


Use filter to remove sites with missing counts.

Find rows that have NA and sites (levels) that have them.

```{r}
(na_rows <- filter(roseate_tidy, is.na(Population)))
(na_rows$Sites <- factor(na_rows$Sites))
(nullsites <-  levels(na_rows$Sites))

```

```{r}
filter(roseate_tidy, Sites != nullsites)
```

Data with sites with all readings.

```{r}
(roseate_new <- filter(roseate_tidy, Sites != "AL" & Sites != "BL" & Sites != "JL"))
```

Make a line plot of population size over time.

Change the axis labels as appropriate (you have to start thinking about what is appropriate).

```{r}
roseate_new %>% 
ggplot() +
  geom_line(mapping = aes(x = year, y = Population, color = Sites)) +
  labs(x = "Years")
```



Which population(s) obviously increased in size between 1992 and 2016? Which population(s) obviously decreased in size during that time?

Populations that increased in size-
AR,
Populations that increased in size-

Some lines have breaks in them. That is, they are not continuous across all years. Why?





###### New start

```{r}
(roseate_tidy <- roseate_data %>% 
  pivot_longer(c(AR, AL, AC, BR, BL,BC, JR, JL, JC ), names_to = "Sites", values_to = "Population"))
```

Use filter to remove sites with missing counts.

```{r}
(roseate_new <- filter(roseate_tidy, !is.na(Population)))
  
```

Make a line plot of population size over time.
Change the axis labels as appropriate (you have to start thinking about what is appropriate).

```{r}
roseate_new %>% 
  ggplot() +
  geom_line(mapping = aes(x = year, y = Population, color = Sites)) +
  labs(x = "Years")
```


Which population(s) obviously increased in size between 1992 and 2016? Which population(s) obviously decreased in size during that time?

Populations that increased in size-
AR, AC, AL,BC, BL, JC, JL
Populations that decreased in size-
JR, BR

Some lines have breaks in them. That is, they are not continuous across all years. Why?

Some data was missing (NA) so it was not plotted, hence why the lines are not continuous. 

#########################

#### Blacklip Abalone

Chunk 1: Import, remove the first column, then make a boxplot of height differences among the three types.

```{r}
abalone_data <- read_csv(here("data", "abalone.csv"))
```

Tidy version (without first column)

```{r}
abalone_data <- select(abalone_data, Type:Rings) 
```

Boxplot of height differences among the three types.

```{r}
abalone_data %>% 
  ggplot() +
  geom_boxplot(aes(x = Type, y = Height))
```

The order of increasing height is - I< M< F. M and F each have an outlier. 


Chunk 2: The boxplot for height shows a female and a male outlier. Perhaps the samples contained two very large, old individuals. Make a scatterplot to see if height appears to correlate with rings. Rings is a measure used to estimate age. Based on the graph, are the extraordinarily large individuals really old individuals?

```{r}
abalone_data %>% 
  ggplot() +
  geom_point(aes(x = Rings, y = Height, color = Type))
```

No, the two individuals are outliers. They are not the oldest and have less than 10 rings.


Chunk 3: Let’s assume the outliers are coding errors so remove them by filtering. Filter the data to remove the two large individuals. Change Type to an ordered factor. Immatures must be first, as that makes sense in terms of age. The order of female and male after immature is up to you. Then, redo the scatterplot that you just made with the newly wrangled data.

What patterns emerge? Which type is the largest? Are all females and males larger than immatures?

```{r}
abalone_filter <- filter(abalone_data, Height < 0.4)
abalone_filter$Type <- factor(abalone_filter$Type, ordered = TRUE, levels = c("I", "M", "F"))

abalone_filter %>% 
  ggplot() +
  geom_point(aes(x = Rings, y = Height, color = Type))
```

Immatures on an average are smaller than M and F . Male and Females largely overlap. No, not all females and males are larger than immatures.


Chunk 4: Are there really immatures with more than five rings with zero height? Srsly? - Print the records of the individuals with zero height. - Most likely, the two zero height values are mistakes made during data recording. This time, instead of filtering them, assign NA (missing data) to those two records. Repolot the data to ensure the two observations are not included in the graph.

Individuals with 0 height

```{r}
filter(abalone_filter, Height == 0)
```


```{r}
abalone_filter[, "Height"][abalone_filter[, "Height"] == 0] <- NA
abalone_filter %>% 
  ggplot() +
  geom_point(aes(x = Rings, y = Height, color = Type))
abalone_filter
```

```{r}
subset(abalone_filter, Height == NA)
```


#####################

###################

Chunk 5: Make two scatterplots of your choice, between any two pairs of continuous variables that make sense to show as scatterplots. Color, shape, or both should distinguish the three types.



```{r}
abalone_filter %>% 
  ggplot() + 
    geom_point(mapping = aes(x = Diameter, y = Rings, color = Type))

abalone_filter %>% 
ggplot() +
    geom_point(mapping = aes(x = ShellWeight, y = WholeWeight, color = Type))
  
```

In Rings vs Diameter, the data shows correlation 

???????????????????????

###################################













#### Darters


```{r}
darters_data <- read_fwf(here("data", "darter.txt"), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3,3, 2, 2)), comment = "#")
```

# 1-3 specimen id
# 4-14 species
# 15 riffle: 1 = upstream riffle, 2 = downstream riffle
# 16 age in year
# 17 sex       
# 18-20 length in centimeters
# 21-24 depth in centimeters
# 25-28 velocity in meters per second
# 29-32 temperature Celsius
# 33-35 majsub proportion of major substrate type
# 36-38 minsub proportion of minor substrate type
# 39-40 majtype: s = sand, fg = fine gravel, sg = small gravel, lg = large gravel, c = cobble
# 41-42 mintype: s = sand, fg = fine gravel, sg = small gravel, lg = large gravel, c = cobble, z = no present







