---
title: 'HW07: Data Importing and Wrangling 1'
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Load the packages and file path.

```{r}
library(tidyverse)
library(here)
```


#### Limpets

Requirements:

Data file: limpets.csv.

Are the data tidy?
Yes, all three columns are unique variables- Length, Site, Protected- each observation is a row.

Use the col_types argument. The three column types are numeric, character, and character.

```{r}
limpets_data <- read_csv(here("data", "limpets.csv"), col_types = "ncc")
```

Make the Sites column an ordered factor with these levels:
PBL, KNRM, VBG, WP, PF, DP, SIO, CTZ1, CTZ2, CNM

```{r}
limpets_data$Site <- factor(limpets_data$Site, ordered = TRUE, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"))
```

PBL is the northern most site (Pebble Beach). CNM is the southernmost site (Cabrillo National Monument).

Make a boxplot of length (in millimeters) for each site, colored by protected status.

Change the axis labels so they begin with capital letters. The y-axis should include the unit of measurement in parentheses.

Note: Both axes labels already begin with capital letters. 


```{r}
limpets_data %>% 
  ggplot() +
  geom_boxplot(aes(x = Site, y = Length, color = Protected)) +
  labs(y = "Length(in millimeters)")
```

In your description, tell which two sites have outliers and whether the protected sites tend to have larger or smaller limpets.

The two sites that have outliers are WP and DP. Protected sites tend to have larger limpets than unprotected ones.






#### Roseate Terns

Data: roseate_terns.txt

Are the data tidy?
No, sites are individual variables (columns) not one variable. Also, one observation is not one row.

```{r}
(roseate_tidy <- roseate_data %>% 
  pivot_longer(c(AR, AL, AC, BR, BL,BC, JR, JL, JC ), names_to = "Sites", values_to = "Population"))
```
roseate_tidy is tidy data.


Use filter to remove sites with missing counts.

```{r}
(roseate_new <- filter(roseate_tidy, !is.na(Population)))
  
```

Make a line plot of population size over time.
Change the axis labels as appropriate (you have to start thinking about what is appropriate).

```{r}
roseate_new %>% 
  ggplot() +
  geom_line(mapping = aes(x = year, y = Population, color = Sites)) +
  labs(x = "Years")
```


Which population(s) obviously increased in size between 1992 and 2016? Which population(s) obviously decreased in size during that time?

Populations that increased in size-
AR, AC, AL,BC, BL, JC, JL
Populations that decreased in size-
JR, BR

Some lines have breaks in them. That is, they are not continuous across all years. Why?

Some observations are missing (NA) so it was not plotted, hence why the lines are not continuous. 



#### Blacklip Abalone

Chunk 1: Import, remove the first column, then make a boxplot of height differences among the three types.

```{r}
abalone_data <- read_csv(here("data", "abalone.csv"))
```

Tidy version (without first column)

```{r}
(abalone_data <- select(abalone_data, Type:Rings)) 
```

Boxplot of height differences among the three types.

```{r}
abalone_data %>% 
  ggplot() +
  geom_boxplot(aes(x = Type, y = Height))
```

The order of increasing height is - I< M< F. M and F each have an outlier. 


Chunk 2: The boxplot for height shows a female and a male outlier. Perhaps the samples contained two very large, old individuals. Make a scatterplot to see if height appears to correlate with rings. Rings is a measure used to estimate age. Based on the graph, are the extraordinarily large individuals really old individuals?

```{r}
abalone_data %>% 
  ggplot() +
  geom_point(aes(x = Rings, y = Height, color = Type))
```

No, the two individuals are outliers. They are not the oldest and have less than 10 rings.


Chunk 3: Let’s assume the outliers are coding errors so remove them by filtering. Filter the data to remove the two large individuals. Change Type to an ordered factor. Immatures must be first, as that makes sense in terms of age. The order of female and male after immature is up to you. Then, redo the scatterplot that you just made with the newly wrangled data.

What patterns emerge? Which type is the largest? Are all females and males larger than immatures?

```{r}
abalone_filter <- filter(abalone_data, Height < 0.4)
abalone_filter$Type <- factor(abalone_filter$Type, ordered = TRUE, levels = c("I", "M", "F"))

abalone_filter %>% 
  ggplot() +
  geom_point(aes(x = Rings, y = Height, color = Type))
```

Immatures on an average are smaller than M and F . Male and Females largely overlap.It is hard to tell from th escatterplot which type is the largest. No, not all females and males are larger than immatures.


Chunk 4: Are there really immatures with more than five rings with zero height? Srsly? - Print the records of the individuals with zero height. - Most likely, the two zero height values are mistakes made during data recording. This time, instead of filtering them, assign NA (missing data) to those two records. Repolot the data to ensure the two observations are not included in the graph.

Individuals with 0 height

```{r}
filter(abalone_filter, Height == 0)
```


```{r}
abalone_filter[abalone_filter$Height == 0, 'Height'] <- NA
```


```{r}
abalone_filter %>% 
  ggplot() +
  geom_point(aes(x = Rings, y = Height, color = Type))
abalone_filter
```
The two observations are not included.


Chunk 5: Make two scatterplots of your choice, between any two pairs of continuous variables that make sense to show as scatterplots. Color, shape, or both should distinguish the three types.



```{r}
abalone_filter %>% 
  ggplot() + 
    geom_point(mapping = aes(x = Diameter, y = Rings, color = Type)) 

abalone_filter %>% 
ggplot() +
    geom_point(mapping = aes(x = ShellWeight, y = WholeWeight, color = Type))+
labs(x = "Shell Weight", "Whole weight")
  
```

In Rings vs Diameter, the data shows I has less rings and diameter than M and F (which overlap for most of observations).

In Whole Weight vs Shell Weight I has less Whole Weight and Shell Weight compared to M and F but here the data is less scattered and the overlap between I and the others is less compared to that of the previous graph.   


#### Darters

Data: darters.txt

```{r}
darters_data <- read_fwf(here("data", "darter.txt"), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2)), comment = "#")
```

# 1-3 specimen id
# 4-14 species
# 15 riffle: 1 = upstream riffle, 2 = downstream riffle
# 16 age in year
# 17 sex       
# 18-20 length in centimeters
# 21-24 depth in centimeters
# 25-28 velocity in meters per second
# 29-32 temperature Celsius
# 33-35 majsub proportion of major substrate type
# 36-38 minsub proportion of minor substrate type
# 39-40 majtype: s = sand, fg = fine gravel, sg = small gravel, lg = large gravel, c = cobble
# 41-42 mintype: s = sand, fg = fine gravel, sg = small gravel, lg = large gravel, c = cobble, z = no present

Data: darters.txt
Column names and widths are included in the file. You can use whatever column names you want but adjust accordingly for info below.
Make riffle an unordered factor with levels 1 and 2.

```{r}
colnames(darters_data) <-  c("ID", "Species", "Riffle", "Age", "Sex", "Length", "Depth", "Velocity", "Temperature", "majsub", "minsub", "majtype", "mintype")

darters_data$Riffle <- factor(darters_data$Riffle)
```

Make major_type an ordered factor with levels s, fg, sg, lg, c

```{r}
darters_data$majtype <- factor(darters_data$majtype, ordered = TRUE, levels = c("s", "fg", "sg", "lg", "c"))
```

Do these four steps together with the pipe.

Filter to remove rows with “zonale” and “tetrazonum”.
Remove mintype and minsub columns.
Rename majsub and majtype to major_substrate and major_type, respectively.
Arrange the data by id.


```{r}
(darters_plot <- darters_data %>% 
  filter(Species == "blennioides" | Species == "caeruleum" | Species == "spectabile") %>% 
  select(ID:majsub, majtype) %>% 
  rename(major_substrate = majsub, major_type =  majtype) %>% 
  arrange(ID))
```

Plot 1 Plot length as a function of depth. Map species to color and shape. What differences do you see between the two riffles?

```{r}
darters_plot %>% 
  ggplot() +
  geom_point(mapping = aes(x = Depth, y = Length, color = Species, shape = Species)) +
  facet_wrap(~ Riffle)
```

The upstream riffle(1) seems to have less *blennioides* than the downstream riffle(2). 2 has the longest length observations. 1 has more observations at greater depths.

Plot 2: Make a boxplot of length for each of the three species. Which riffle shows the greatest number of outliers?

```{r}
darters_plot %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Species, y = Length, color = Species, shape = Species)) +
  facet_wrap(~ Riffle)
```

Riffle 2 (downstream) shows the greatest number of outliers.

Plot 3: Make a boxplot of length for major substrate types for each species for each riffle. This will actually be six plots in one! To do this, use facet_grid(species ~ riffle) in place of facet_wrap(). How does the plot change if you switch the order of the argument (riffle ~ species) in the facet_grid() function?

```{r}
darters_plot %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = major_substrate, y = Length, color = Species, shape = Species)) +
  facet_grid(Species~ Riffle)
```
```{r}
darters_plot %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = major_substrate, y = Length, color = Species, shape = Species)) +
  facet_grid(Riffle ~ Species)
```
Whichever varaible is before "~" in facet_grid() is split on the y axis (row wise) and the one that comes after "~" gets split column wise.
Average length of each species varies in each riffle. *caeruleum* and *spectabile* approximately the same length (with each other) but vary greatly in length between riffles.  They are longer in 1 and shorter in 2 compared to *blennioides*.



#########
