---
title: 'HW07: Data Importing and Wrangling 1'
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Load the packages and file path.

```{r}
library(tidyverse)
library(here)
```


#### Limpets

Requirements:

Data file: limpets.csv.

Are the data tidy?
Yes, all three columns are unique variables- Length, Site, Protected- each observation is a row.

Use the col_types argument. The three column types are numeric, character, and character.

```{r}
limpets_data <- read_csv(here("data", "limpets.csv"), col_types = "ncc")
```

Make the Sites column an ordered factor with these levels:
PBL, KNRM, VBG, WP, PF, DP, SIO, CTZ1, CTZ2, CNM

```{r}
limpets_data$Site <- factor(limpets_data$Site, ordered = TRUE, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"))
```

PBL is the northern most site (Pebble Beach). CNM is the southernmost site (Cabrillo National Monument).

Make a boxplot of length (in millimeters) for each site, colored by protected status.

Change the axis labels so they begin with capital letters. The y-axis should include the unit of measurement in parentheses.

Note: Both axes labels already begin with capital letters. 


```{r}
limpets_data %>% 
  ggplot() +
  geom_boxplot(aes(x = Site, y = Length, color = Protected)) +
  labs(y = "Length(in millimeters)")
```

In your description, tell which two sites have outliers and whether the protected sites tend to have larger or smaller limpets.

The two sites that have outliers are WP and DP. Protected sites tend to have larger limpets than unprotected ones.

#### Roseate Terns

Data: roseate_terns.txt

```{r}
roseate_data <- read_tsv(here("data", "roseate_terns.txt"))
```

Are the data tidy?
No, the sites are individual columns rather than being under one variable of "Sites".

```{r}
(roseate_tidy <- roseate_data %>% 
  pivot_longer(c(AR, AL, AC, BR, BL,BC, JR, JL, JC ), names_to = "Sites", values_to = "Population"))
```

Variables "Population" and "Sites" assigned. Data tidied. New data is 'roseate_tidy' 

Use filter to remove sites with missing counts.

```{r}
na_rows <- filter(roseate_tidy, is.na(Population)) 

```


```{r}
filter(roseate_data$Population != NA)
```


Make a line plot of population size over time.
Change the axis labels as appropriate (you have to start thinking about what is appropriate).
Which population(s) obviously increased in size between 1992 and 2016? Which population(s) obviously decreased in size during that time?
Some lines have breaks in them. That is, they are not continuous across all years. Why?
















