---
title: "HW 10: Data Visualization I"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
#### Load libraries

```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```


***
#### Import and wrangle

```{r}
sparrow_raw <- read_tsv(here("data", "sparrows.txt"))
```

Create two dummy variables.

```{r}
sparrows <- sparrow_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
```


***
#### Figure 2: outliers

Make boxplot similar to Figure 2.

```{r}
(plot_box <- sparrows %>%
  ggplot(aes(x = box_group,
             y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL,
       y = "Wing length (mm)"))
```

Assess full range of values can be assessed with a Cleveland dot plot.

```{r}
(plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "Order of the data"))
```

Recreate exact Figure 2 using patchwork.

```{r}
plot_box + plot_cleveland
```

Align "Wing length (mm)" on y axis.

```{r}
plot_box + plot_cleveland + coord_flip()
```


***
#### Figure 3: outliers

Write a single chunk of code that makes the six graphs shown in Figure 3 and lays them out as shown.

```{r}
p_culmen <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

(p_culmen | p_nalospi | p_wt) /
(p_wingcrd | p_tarsus | p_head)

```

Question: Which three variables appear to have outlier data?
The nalopsi, tarsus and head variable appear to have outliers.


***
#### Figure 5: distribution

Create figure 5.

```{r}
# Histogram
(h1 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14,28, by = 2)))
```

Filters out all months except 6, 7, or 8, then mutate the numbers to the month names.

```{r}
# Faceted Histogram
(h2 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>%
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1))
```

Final layout
```{r}
h1 + h2
```

*Honors*

Recreate this figure using density plots. However, modify the right panel so that all three months are overlaid on the same plot. Go back to the notes and follow the example code for guidance.



```{r}
d1 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  ggplot() +
  geom_density(aes(x = wt)) +
  scale_x_continuous(breaks = seq(14,28, by = 2))

d2 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>%
  ggplot() +
  geom_density(aes(x = wt, fill = Month),
                 alpha = 0.42) +
  scale_x_continuous(breaks = seq(14, 28, by = 2))
d1 + d2
```

***
#### Figure 10: Relationships

```{r}
sparrows %>% ggscatmat(columns = 1:7)
```


***
#### Figure 11: interactions

```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>%
  filter(Month %in% 5:9,
         Sex != 0) %>%
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>%
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month) 
```


### END

