---
title: "HW 10: Data Visualization I"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
#### Load libraries

```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```


***
#### Import and wrangle

```{r}
sparrow_raw <- read_tsv(here("data", "sparrows.txt"))
```

Create two dummy variables.

```{r}
sparrows <- sparrow_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
```


***
#### Figure 2: outliers

Make boxplot similar to Figure 2.

```{r}
(plot_box <- sparrows %>%
  ggplot(aes(x = box_group,
             y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL,
       y = "Wing length (mm)"))
```

Assess full range of values can be assessed with a Cleveland dot plot.

```{r}
(plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "Order of the data"))
```

Recreate exact Figure 2 using patchwork.

```{r}
plot_box + plot_cleveland
```

Align "Wing length (mm)" on y axis.

```{r}
plot_box + plot_cleveland + coord_flip()
```


***
#### Figure 3: outliers

Write a single chunk of code that makes the six graphs shown in Figure 3 and lays them out as shown.

```{r}
p_culmen <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

(p_culmen | p_nalospi | p_wt) /
(p_wingcrd | p_tarsus | p_head)

```

Question: Which three variables appear to have outlier data?
The nalopsi, tarsus and head variable appear to have outliers.


















