---
title: "HW 14: Maps"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---




###############
Change the Knit directory to “Project Directory”???
#############



***
#### Load libraries


```{r}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```


***
#### Part 1: Outline maps

**Example: States and counties**

*U.S. map*

Access specific regions (states, in this case) and store the results in three data frames.

```{r}
# Outline of the 48 contiguous states
usa <- map_data(map = "state")

# Outline of just Missouri
us_missouri <- map_data(map = "state",
                        region = "missouri")

# Outline of Alabama, Florida, Mississippi, and North Carolina
us_other <- subset(x = usa, 
                   subset = region %in% c("alabama",
                                          "florida",
                                          "mississippi",
                                          "north carolina"))
```

Make a map with layers from above data.

The first layer has the four non-Missouri states.

```{r}
ggplot() +
  geom_polygon(data = us_other,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray")
```

Add Missouri, filled with black to emphasize it over the other four states.

```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "black")
```

Add outline of 48 states.

```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long,
                   y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black")
```

Replace the default theme with theme_map(). Fix the longitude 1.3 times longer than the latitude. Add the filled states first so that the black outlines of all states are overlaid on top. Name it us_map.

```{r}
us_map <- ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray") +
  geom_polygon(data = us_missouri, 
               aes(x = long,
                   y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black") +
  theme_map() +
  coord_fixed(1.3)

us_map
```


*Missouri map*
Highlight Scott County to show the county where the ant species was collected for the first time ever. Add the Scott County layer first, then the outline of all the counties.

```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri,
                   subregion %in% c("scott"))

mo_map <- ggplot() +
  geom_polygon( data = missouri,
                aes(x = long,
                    y = lat,
                    group = group),
                fill = NA,
                color = "black") +
  geom_polygon( data = mo_scott,
                aes(x = long,
                    y = lat,
                    group = group),
                fill = "black") +
  theme_map() +
  coord_fixed(1.3)

mo_map
```

Put both maps in one row and make us map 1.5 times wider that missouri.

```{r}
us_map + mo_map +
  plot_layout(ncol = 2,
              widths = c(1.5, 1))
```


**Outline maps: your turn**































