---
title: 'HW06 Part 1: Complete the sections'
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

Install package

```{r}
library(nycflights13)
```


## Chapter 10: Tibbles

Read [R4ds Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html), sections 1-3.

### 10.1: Introduction

Load the tidyverse package. 

```{r echo=FALSE}
library(tidyverse)
```


### 10.2: Creating tibbles

Enter your code chunks for Section 10.2 here.

Describe what each chunk code does. 

Coerce a data frame to a tibble

```{r}
as_tibble(iris)
```

Create a new tibble from individual vectors

```{r}
tibble(
  x = 1:5,
  y = 1,
  z = x^2 + y
)
```

Give non-syntactic names to columns using tibble and view it

```{r}
tb <- tibble(
  `:)` = "smile",
  ` ` = "space",
  `2000` = "number"
)
tb
```

Create a tibble with tribble() i.e. customized data entry in code 

```{r}
tribble(
  ~x, ~y, ~z,
  #--/--/----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

### 10.3: Tibbles vs data.frame

Enter your code chunks for Section 10.3 here.

Describe what each chunk code does. 

Created a tibble to view its refined print method

```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

View a large data frame in default

```{r}
nycflights13::flights %>%
  print(n = 10, width = Inf)
```

Get a scrollable view of a complete dataset.

```{r}
nycflights13::flights %>% 
  View()
```

Extract variables in tibble

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

# Extract by name
df$x
df[["x"]]

# Extract by position
df[[1]]
```

Use extraction in pipes

```{r}
df %>% .$x
df %>% .[["x"]]
```

### 10.4: Not required

#### Section 10.5 Questions

Answer the questions *completely.* Use code chunks, text, or both, as necessary.

**1:** How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame). Identify at least two ways to tell if an object is a tibble. *Hint:* What does `as_tibble()` do? What does `class()` do? What does `str()` do?

When viewed as data frame

```{r}
iris
```

View as tibble

```{r}
irist <- as_tibble(iris)
irist
```
When viewing data frame it shows all rows whereas when viewing a tibble it shows only the first 10 rows with the type of data each variable has under it. 
When using class() on a data frame it shows "data.frame", when using it on tibble it shows-
"tbl_df"     "tbl"        "data.frame"



**2:** Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

As tibble

```{r}
df1 <- as_tibble(df)
df1$x
df1[, "xyz"]
df1[, c("abc", "xyz")]
```
The difference is that when unfinished name of column is added after '$' in dataframe it shows the result with its closest match whereas tibble shows NULL. This can cause mistakes as it may match to incorrect variable. 
Also,class(df[, "xyz"]) is "character" but class(df1[, "xyz"]) is-
[1] "tbl_df"     "tbl"        "data.frame"
i.e. tibble retains its type even when result is a single outcome.



## Chapter 11: Importing data

Read [R4ds Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html), sections 1, 2, and 5.

### 11.1 Introduction

Nothing to do here unless you took a break and need to reload `tidyverse`.

### 11.2 Getting started.

Do *not* run the first code chunk of this section, which begins with `heights <- read_csv("data/heights.csv")`. You do not have that data file so the code will not run.

Enter and run the remaining chunks in this section.

Supply an inline csv file.

```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```

Skip a few lines of metadata at the top of the file.

```{r}
read_csv("The first line of metadata
The second line of metadata
x,y,z
1,2,3", skip = 2)

read_csv("# A comment I want to skip
x,y,z
1,2,3", comment = "#")
```

Make the data not have column names

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

Make a character vector the column names

```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

Make value '.' stand for missing values in your data

```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```


#### 11.2 Questions

**1:** What function would you use to read a file where fields were separated with "|"?
read_delim() should be used. Here delim = "|".


**2:** (This question is modified from the text.) Finish the two lines of `read_delim` code so that the first one would read a comma-separated file and the second would read a tab-separated file. You only need to worry about the delimiter. Do not worry about other arguments. Replace the dots in each line with the rest of your code. 

# Comma-separated
`file <- read_delim("file.csv", delim = ",")`

# Tab-separated
`file <- read_delim("file.csv", delim = "\t")`


**3:** What are the two most important arguments to `read_fwf()`? Why?
The two common arguments are fwf_widths() and fwf_positions() .Widths of files can be specified with fwf_widths() and their position with fwf_positions().

Also, read_table() reads a common variation of fixed width files where columns are separated by white space.

**4:** Skip this question


**5: ** Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

1.A column is missing in the first row. Only two columns made based on first row, values under third column discarded

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
```

2.Uneven number of columns. Number of columns set by first row. Missing value filled by NA. Extra value in row three left out.

```{r}
read_csv("a,b,c\n1,2\n1,2,3,4")
```

3.Only one output in second row for 2 columns. Also, there is an open quote infront of 1. Value under second column in second row filled as NA
```{r}
read_csv("a,b\n\"1")
```

4.Nothing is wrong. (Though the first and third row have similar values). The variables are character vectors.
```{r}
read_csv("a,b\n1,2\na,b")
```
5.Semicolons used instead of commas. Makes one column and substitutes both values under it as one. Could use read_csv2. 
```{r}
read_csv("a;b\n1;3")
```

### 11.3 and 11.4: Not required

### 11.5: Writing to a file

Just read this section. You may find it helpful in the future to save a data file to your hard drive. It is basically the same format as reading a file, except that you must specify the data object to save, in addition to the path and file name.

### 11.6 Not required

## Chapter 18: Pipes

Read [R4ds Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html), sections 1-3.

Nothing to do otherwise for this chapter. Is this easy or what?

**Note:** Trying using pipes for all of the remaining examples. That will help you understand them.

## Chapter 12: Tidy Data

Read [R4ds Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html), sections 1-3, 7. 

### 12.1 Introduction

Nothing to do here unless you took a break and need to reload the `tidyverse.`

### 12.2 Tidy data

Study Figure 12.1 and relate the diagram to the three rules listed just above them. Relate that back to the example I gave you in the notes. Bear this in mind as you make data tidy in the second part of this assignment.

You do not have to run any of the examples in this section.

### 12.3

Read from the start of the chapter through 12.3.1. Run the examples in section 12.3.1 (Longer), including the example with `left_join()`. We'll cover joins later. Skip section 12.3.2.

*12.3.1*

Making Table 4a

```{r}
table4a <- read_csv("country,1999,2000
Afghanistan,745,2666
Brazil,37737,80488
China,212258,213766")
table4a
```

Make the column names variables not values

```{r}
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```

Making table 4b

```{r}
table4b <- read_csv("country,1999,2000
Afghanistan,19987071,20595360
Brazil,172006362,174504898
China,1272915272,1280428583")
table4b
```

pivot_longer table 4b with values_to = "population"

```{r}
table4b %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
```

Combine table 4a and 4b into a single tibble

```{r}
tidy4a <- table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
tidy4b <- table4b %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
left_join(tidy4a, tidy4b)
```


#### 12.3.3 Questions

**2:** Why does this code fail? Fix it so it works.

```{r error=TRUE}
table4a %>% 
  pivot_longer(c(1999,2000), names_to = "year", values_to = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```

It fails because 1999 and 200 are non-syntactic names (because they don’t start with a letter) so we have to surround them in backticks.

Fixed code-
```{r}
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```


Skip the rest of Chapter 12. On to the last chapter.


## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data Transformation](https://r4ds.had.co.nz/transform.html), sections 1-4.

Time to [get small.](https://www.youtube.com/watch?v=GOrdzCHnpw4) 

### 5.1: Introduction

Load the necessary libraries. As usual, type the examples into and run the code chunks.

The tidyverse and nycflights13 packages already loaded.

View head of data

```{r}
flights
```


### 5.2: Filter rows with `filter()`

Study Figure 5.1 carefully. Once you learn the `&`, `|`, and `!` logic, you will find them to be very powerful tools.

Select all flights on January 1st

```{r}
filter(flights, month == 1, day == 1)
```

Save the results of filter

```{r}
jan1 <- filter(flights, month == 1, day == 1)
```

Save and view results of filter for another date

```{r}
(dec24 <- filter(flights, month == 12, day == 25))
```

*5.2.1 Comparisons*

Use comparison operators.

```{r error= TRUE}
filter(flights, month == 1)
```
Corrected
```{r}
filter(flights, month == 1)
```

Floating point numbers

```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```

Using near() to see in approximation

```{r}
near(sqrt(2) ^ 2, 2)
near(1 / 49 * 49, 1)
```

*5.2.2 Logical operators*

All flights that departed in November or December

```{r}
filter(flights, month == 11 | month == 12)
```
Rewrite the above using short hand. {"()" added to code to view it}

```{r}
(nov_dec <- filter(flights, month %in% c(11, 12)))
```

Flights that weren’t delayed (on arrival or departure) by more than two hours

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

*5.2.3 Missing values*

Operations involving an unknown value

```{r}
NA > 5
10 == NA
NA + 10
NA / 2
```

```{r}
NA == NA
```

Expression with context

```{r}
# Let x be Mary's age. We don't know how old she is.
x <- NA

# Let y be John's age. We don't know how old he is.
y <- NA

# Are John and Mary the same age?
x == y

# NA in answer implies that we do not know. 
```

Determine if a value is missing

```{r}
is.na(x)
```

Preserve TRUE and NA values

```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
filter(df, is.na(x) | x > 1)
```


#### 5.2 Questions

**1.1:** Find all flights with a delay of 2 hours or more.

```{r}
filter(flights, dep_delay >= 120 | arr_delay >= 120)
```
Note: The question in text book states "arrival delay" so its code would be-
```{r}
filter(flights, arr_delay >= 120)
```


**1.2:** Flew to Houston (IAH or HOU)

```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

**1.3:** Were operated by United (UA), American (AA), or Delta (DL).

```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

**1.4:** Departed in summer (July, August, and September).
```{r}
filter(flights, month %in% c(7:9))
```

**1.5:** Arrived more than two hours late, but didn’t leave late.
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```

**1.6:** Were delayed by at least an hour, but made up over 30 minutes in flight. This is a tricky one. Do your best.
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30 )
```

**1.7:** Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time %in% c(0000:0600, 2400))
```

**2:** Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?
This is a shortcut for x >= left & x <= right.

```{r}
filter(flights, between(month,7,9))
```

**3:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

```{r}
nrow(filter(flights, is.na(dep_time)))
```
Number of flights with a missing dep_time is 8255.
Other variables such as dep_delay, arr_time, arr_delay and air_time are also missing. These rows represent the cancelled flights.

**4:** Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

**Note:** For some context, see [this thread](https://blog.revolutionanalytics.com/2016/07/understanding-na-in-r.html)

```{r}
NA ^ 0
NA | TRUE
FALSE & NA
```
Any number (even unknown) raised to power 0 is equal to 1. 
In the last two expression NA has a logical value. In NA or TRUE, atleast one value has to be TRUE which it is irrespective of what value NA has. Anything 'and' "FALSE" is always false irrespective of what this "anything" is.


### 5.3 Arrange with `arrange()`

Rearrange columns in 'flights'

```{r}
arrange(flights, year, month, day)
```

Use desc() to re-order by a column in descending order

```{r}
arrange(flights, desc(dep_delay))
```

Sort missing values

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df, desc(x))
```
Note- Default of arrange() in increasing order if only one variable mentioned.

#### 5.3 Questions

**1:** How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()). **Note:** This one should still have the earliest departure dates after the `NA`s. *Hint:* What does `desc()` do?

```{r}
arrange(flights, desc(is.na(dep_time)), year, month, day, sched_dep_time)
```


**2:** Sort flights to find the most delayed flights. Find the flights that left earliest. 

This question is asking for the flights that were most delayed (left latest after scheduled departure time) and least delayed (left ahead of scheduled time).

Flights that were most delayed

```{r}
arrange(flights, desc(dep_delay))
```

Flights that were most ahead in time for departure

```{r}
arrange(flights, dep_delay)
```


**3:** Sort flights to find the fastest flights. Interpret fastest to mean shortest time in the air.

```{r}
arrange(flights, air_time)
```


*Optional challenge:* fastest flight could refer to fastest air speed. Speed is measured in miles per hour but time is minutes. Arrange the data by fastest air speed.

```{r}
arrange(flights, desc(distance / air_time))
```

**4:** Which flights travelled the longest? Which travelled the shortest?

Longest flights
```{r}
arrange(flights, desc(distance))
```

Shortest flights

```{r}
arrange(flights, distance)
```


### 5.4 Select columns with `select()`

Select columns by name

```{r}
select(flights, year, month, day)
```

Select all columns between year and day (inclusive)

```{r}
select(flights, year:day)
```

Select all columns except those from year to day (inclusive)

```{r}
select(flights, -(year:day))
```

Rename a variable using rename()

```{r}
rename(flights, tail_num = tailnum)
```

Move handful of variables to the start of the data frame.

```{r}
select(flights, time_hour, air_time, everything())
```


#### 5.4 Questions

**1:** Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights. Find at least three ways.

Using variable names of columns.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

Using column names
```{r}
select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")
```

Using column numbers

```{r}
select(flights, 4, 6, 7, 9)
```


**2:** What happens if you include the name of a variable multiple times in a `select()` call?

```{r}
select(flights, dep_time, dep_delay, dep_time)
```
It only registers the first time it was mentioned and does not duplicate. It does not show an error 

**3:** What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```
'one_of' selects columns from their character names (in a vector) rather than their unquoted variable names.
'vars' is a character vector with column names one can select using 'one_of'

```{r}
select(flights, one_of(vars))
```


**4:** Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME"))
```

Yes, most since most functions are case sensitive. The select helpers ignore case by default.
This can be changed by-

```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

### THE END