---
title: "HW 13: Strings and things"
author: "Afreen Alam"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
#### Load libraries

```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```

***
#### Part 1: DNA strings

**Import the data**

```{r}
beta_gliadin_raw <- read_lines(here("data", "beta_gliadin.txt"), skip = 1)
head(beta_gliadin_data)
```


**From many strings to one**


Replace all of the whitespace with nothing.

Replace all of the digits with nothing.

```{r}
(beta_data <- beta_gliadin_raw %>%
  str_replace_all(c(" " = "", "\\d" = "", "\\t" = "")))
```

Use str_c to collapse the strings. Print the variable to ensure you have just one string. Here are the first 100 characters.

```{r}
(beta_data_c <- beta_data %>%
   str_c(collapse = "")
   )
```

Convert it to upper case with str_to_upper().

```{r}
(beta_data_final <- beta_data_c %>%
   str_to_upper()
   )
```

 How long is the total DNA sequence? You should get 3310 characters.
 
```{r}
str_length(beta_data_final)
```

**Sequence analysis: find the promoter region**

Use str_count() and regex to count the number of IUPAC letters in the sequence that are not A, C, G, or T.
(Ans - None)

```{r}
str_count(beta_data_final, "[^ATCG]")
```

Count how many possible start codons are in the sequence.

```{r}
str_count(beta_data_final, "ATG")
```
Count how many possible stop codons are located in the sequence. For full credit, I’ll be more impressed if you search for both stop codons with a single regular expression. Hint: |

```{r}
str_count(beta_data_final, "(TGA|TAA)")
```



































































